---
title: "Risk Parity Plots"
author: "Jack Ye"
date: "July 25, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DataComputing)
```

# weights change
```{r warning=F}
read.csv('~/Desktop/innovation/riskp/weights2.csv', header=TRUE, sep=',') %>%
  gather(time, weight, -Ticket) %>%
  mutate(time=as.numeric(substr(time, 2, length(time)+1))) %>%
  ggplot(aes(x=time, y=weight, col=Ticket)) + geom_line()
```
\newpage

# daily return, before brexit
general trend up, similar volatility as the market

```{r warning=F}
read.csv('~/Desktop/innovation/riskp/dailyPercentage.csv', header=TRUE, sep=',') %>%
  gather(time, weight, -Ticket) %>%
  mutate(time=as.numeric(substr(time, 2, length(time)+1))) %>%
  filter(time %% 5 == 0) %>%
  filter(time >= 938) %>%
  ggplot(aes(x=time, y=weight, col=Ticket)) + geom_line()
```
\newpage

# cumulative return, before brexit
general trend up, colsely following the market trend

```{r warning=F}
read.csv('~/Desktop/innovation/riskp/cumulativeReturn.csv', header=TRUE, sep=',') %>%
  gather(time, weight, -Ticket) %>%
  mutate(time=as.numeric(substr(time, 2, length(time)+1))) %>%
  filter(time %% 3 == 0) %>%
  ggplot(aes(x=time, y=weight, col=Ticket)) + geom_line()
```
\newpage

# Brexit daily return
Volatile period, less volatile than the market
```{r warning=F}
read.csv('~/Desktop/innovation/riskp/brexitDaily.csv', header=TRUE, sep=',') %>%
  gather(time, weight, -Ticket) %>%
  mutate(time=as.numeric(substr(time, 2, length(time)+1))) %>%
  filter(time %% 5 == 0) %>%
  ggplot(aes(x=time, y=weight, col=Ticket)) + geom_line()
```
\newpage

# Brexit cumulative return
Investing risk parity portfolio at a volatile period gives higher return
```{r warning=F}
read.csv('~/Desktop/innovation/riskp/brexitCumulative.csv', header=TRUE, sep=',') %>%
  gather(time, weight, -Ticket) %>%
  mutate(time=as.numeric(substr(time, 2, length(time)+1))) %>%
  ggplot(aes(x=time, y=weight, col=Ticket)) + geom_line()
```





